# This File is for exploratory Data Analysis and Cleaning for R Shiny Project
```{r global}

##STARTING WITH EVERYTHING SHINY APP WOULD HAVE FROM GLOBAL SCRIPT

#loading packages
library(lubridate)
library(ggplot2)
library(ggmap)
library(ggthemes)
library(dplyr)
library(data.table)
library(tidyverse)
library(shiny)
library(shinydashboard)
library(maps)
library(DT)
library(leaflet)
theme_set(theme_minimal())
library(vroom) #fast data reading/importing
library(sf) #spatial data
library(tigris) #geojoin
library(htmlwidgets) #interactive map labels

#loading cleaned data sets
#preprocessing of data can be found in the preProc.R file within the Shiny App
crash <- read.csv("./ShinyApp/data/crashes_clean.csv", stringsAsFactors = FALSE) #cleaned crash data involving at least one bicycle between 2012-2022
reqs <- read.csv("./ShinyApp/data/bike_reqs_clean.csv", stringsAsFactors = FALSE) #Safety requests shared by bikers, or shared by drivers and pedestrians regarding bikers

#loading shapefiles for wards
wards <- st_read("./ShinyApp/data/wards/Wards_from_2022.shp")


##Modifying Raw Data
wards <- wards %>% select(ward=WARD,
                          name=NAME
)
```

```{r}
geom_bar()
```


```{r}
trial <- reqs %>%
  group_by(reqtype) %>%
  count(year)
#count(reqs, reqtype)
trial
```
```{r}
crash.injtype <- ordered(crash, levels = c("fatality", "major_injury", "minor_injury", "unknown_injury", "no_injury"))
is.factor(crash.injtype)
```


```{r}
ward_dens <- crash %>%
  group_by(ward) %>%
  summarize(incidents = n())

```
```{r}
ward_dens
```


```{r}
wards_1 <- geo_join(wards, ward_dens,
                   'name', 'ward',
                   how = 'inner')
```
```{r}
saveRDS(wards_1, "./data/wards_1.RDS")
```


group_by(.data, ..., .add = FALSE, .drop = group_by_drop_default(.data))

```{r}
wards_1 %>%
  ggplot(aes(x=as.numeric(incidents))) +
  geom_histogram(bins=20, fill ='#69b3a2', color="white")
```

```{r}
##MAKE INTERACTIVE MAP
labels <- sprintf(
  "<strong>%s</strong><br/>%g Total incidents 2012-2022",
  wards_1$name, wards_1$incidents) %>%
  lapply(htmltools::HTML)

pal <-colorBin(palette = "OrRd", 9, domain = wards_1$incidents)

map_interactive <- wards_1 %>%
  st_transform(crs = "+init=epsg:4326") %>%
  leaflet() %>%
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(label= labels,
              stroke = FALSE,
              smoothFactor = .5,
              opacity = 1,
              fillOpacity = 0.7,
              fillColor = ~ pal(incidents),
              highlightOptions = highlightOptions(weight = 5,
                                                  fillOpacity = 1,
                                                  color = "black",
                                                  opacity = 1,
                                                  bringToFront = TRUE))
addLegend("bottomright",
          pal= pal,
          values = ~ incidents,
          title = "Bike Safety Incidents",
          opacity = 0.7)

saveWidget(1, "ward_incident_density.html")
```






```{r}
m <- leaflet() %>% setView(lng = -71.0589, lat = 42.3601, zoom = 12)
m %>% addTiles()
```





































































```{r creating map}
dcmap <- get_stamenmap(
  bbox = c(left = -77.225, bottom = 38.84 , right = -76.8125 , top = 39.000),
  maptype = "terrain",
  crop = TRUE,
  color = "color",
  zoom = 12
)
```
```{r viewing map}
dc <- ggmap(dcmap)
```

```{r}
dc
```
```{r making map + viz + NEED HELP}
#visualizing safety requests
dc +
  geom_point(data = reqs,
             aes(x = lon,
                 y = lat),
             size = .7) +
  theme_map() +
  theme(legend.background = element_blank(),
        legend.position = "bottom")

```

```{r NEEDS HELP}
crash %>% count(ward) %>%
  group_by(injtype)
```
#filtering for incidents of injury
crash = crash %>% filter(inj %in% c("Fatality","Major Injury", "Minor Injury", "Unknown Injury"))

```{r visualizing injury and fatality}

dc +
  geom_point(data = crash,
             aes(x = lon,
                 y = lat,
                 color = injtype),
             size = .7)

```


```{r}

dc +
  geom_point(data = crash,
             aes(x = lon,
                 y = lat,
                 color = fatality),
             size = .7)
```



#getting base map
dcmap <- get_stamenmap(
  bbox = c(left = -77.225, bottom = 38.84 , right = -76.8125 , top = 39.000),
  maptype = "terrain",
  crop = TRUE,
  color = "color",
  zoom = 12
)
#setting base map as dc
dc <- ggmap(dcmap)

```{r}
getColor <- function(crash) {
  sapply(crash$injtype, function(injtype) {
    if(injtype == "no_injury") {
      "green"
    } else if(injtype == "minor_injury" | injtype == "unknown_injury") {
      "yellow"
    } else if(injtype == "major_injury") {
      'Orange'
    } else if(injtype == "fatality"){
      "red"
    } })
}
```
```{r}
getColor(crash)
```
```{r}
icons <- awesomeIcons(
  icon = 'bicycle_outline',
  iconColor = 'black',
  library = 'ion',
  markerColor = getColor(crash)
)
```


```{r}
head(icons)
```
```{r}
r <- leaflet(data = crash) %>% setView(lng = -77.0369, lat = 38.9072, zoom = 12)
r %>% addTiles() %>%
  addAwesomeMarkers(~lon, ~lat, icon = icons, popup = ~as.character(rdate), label = ~as.character(injtype),
             clusterOptions = markerClusterOptions())
```
```{r}
# first 20 quakes
df.20 <- crash[1:20,]

getColor <- function(df.20) {
  sapply(df.20$injtype, function(injtype) {
  if(injtype == 'no_injury') {
    color = "green"
  } else if(injtype <= 'major_injury') {
    color = "orange"
  } else {
    color = "white"
  } })
}

icons <- awesomeIcons(
  icon = 'bicycle',
  iconColor = 'black',
  library = 'ion',
  markerColor = getColor(df.20)
)

leaflet(df.20) %>% addTiles() %>%
  addAwesomeMarkers(~lon, ~lat, icon=icons, label=~as.character(injtype))
```

